\newcommand{\ug}{[xcov-ug]}
\newcommand{\erts}{[ERTS2010]}
\newcommand{\adae}{[AE2009]}
\newcommand{\refdob}{[DO-178B]}
\newcommand{\refdoc}{[DO-178C]}

\section{Referenced documents}
\paragraph*{\adae{}} \ \\
Bordin et al.: Couverture: An Innovative Open Framework for Coverage Analysis
of Safety Critical Applications - Ada User Journal, December 2009.
%\paragraph*{\castseventeen} \ \\
%Certification Authorities Paper 17: Structural coverage of object code
\paragraph*{\refdob{} and [ED-12B]} \ \\
EUROCAE: Software Considerations in Airborne Systems and Equipment
Certification.
\paragraph*{\refdoc{} and [ED-12C]} \ \\
EUROCAE: Software Considerations in Airborne Systems and Equipment
Certification.
\paragraph*{\erts{}} \ \\
Bordin et al: Couverture: An Innovative and Open Coverage Analysis Framework
for Safety-Critical Applications- ERTS2 2010
\paragraph*{\ug{}}
AdaCore: \xcov{} Fundamentals and Users Guide

\section{List of abbreviations}

Below is a list of commonly used abbreviations, with references to documents
where the definitions could be found:

\begin{tabular}{|lll|}
\hline
\B{Abbrev} & \B{Stands for ...} & \B{Definition in} \\ \hline
\stc{} & Statement Coverage & \refdob{} \& \ug{}\\  \hline
\dc{} & Decision Coverage & \refdob{} \& \ug{}\\  \hline
\mcdc{} & Modified Condition/Decision Coverage & \refdob{} \& \ug{}\\  \hline
\tqp{} & Tool Qualification Plan & \\\hline
\tor{} & Tool Operational Requirement & \tqp{}, section \ref{sec:devstd-overview}\\\hline
\tc{} & Test Case & \tqp{}, section \ref{sec:devstd-overview}\\\hline
\end{tabular}

\section{Organization and roles}
%
The development, verification and qualification processes for \xcov{} involve
several teams and parties with established responsibilities:
%
\begin{itemize}
\item%
  The \B{\xcov{} Development Team} contributes to the development of \xcov{},
  including requirements specification, implementation, test cases development
  and test execution.
  %
  This team is in charge of the configuration management of the artifacts it
  produces.
  %
\item%
  The \B{\xcov{} Qualification Team} is responsible for the infrastructure
  supporting the qualification process of \xcov{}.
  % 
  The Qualification Team supports the development team and is also in charge
  of the configuration management of the artifacts it produces.
  %
\item%
  The \B{\xcov{} Quality Assurance Team} is a project-independent team
  responsible to ascertain the expected processes have been put in place.
  %
  The Quality Assurance Team is granted the authority to require specific
  activities to be performed by the \xcov{} Development and Qualification
  Teams.
  %
  This team is also in charge of the configuration management of the
  artifacts it produces (mostly Quality Assurance reports).
  %
\item \textbf{\xcov{} Users} are expected to perform the activities identified
  in section \ref{sec:qualified-use}.
\end{itemize}
