\documentclass {report}
\usepackage{couverture}
\usepackage{color}
\definecolor{light-gray}{gray}{0.85}
\usepackage{listings}
\lstset{backgroundcolor=\color{light-gray}}

\begin{document}
\title{\huge xCov\\Software Quality Assurance Plan\\ \ \\ \large \textbf{Document Version 1.0}}

\maketitle
\tableofcontents

\chapter{Document introduction}

\section{Document purpose}
The intent of the Software Quality Assurance Plan is to describe the process to ascertain the correct application of the Tool Qualification Plan and of the Software Configuration Management Plans.

\section{Authors}
\begin{tabular}{|c|c|c|}
\hline
\textbf{Name} & \textbf{Company} & \textbf{Email} \\ \hline
Matteo Bordin & AdaCore & bordin@adacore.com \\ \hline
Olivier Hainque & AdaCore & hainque@adacore.com \\ \hline
\end{tabular}

\section{Major revision history}
The evolution of this document is automatically tracked by the configuration management system. Here we just provide the major revision history.
\ \\ \\
\begin{tabular}{|c|c|c|}
\hline
\textbf{Version} & \textbf{Date} & \textbf{Comment} \\ \hline
 &  &  \\ \hline
\end{tabular}

\include{doc_introduction}

\chapter{Quality assurance activities}
This sections contains a description of the Quality Assurance activities in terms of objective and output.

\section{Reading of \xcov qualification material: plans}
\begin{itemize}
\item \textbf{objectives:} to assess the compliance with qualification objectives;
\item \textbf{output:} QA Reading report (qa/DDMMYYYY/qa\_plans.doc).
\end{itemize}

\section{Inspection of Tool Operational Requirements (by sampling)}
\begin{itemize}
\item \textbf{objectives:}
\begin{itemize}
\item check the accuracy and completeness with the targeted language feature and coverage metric;
\item check the compliance with the development standard.
\end{itemize}
\item \textbf{output:} QA inspection report (qa/DDMMYYYY/qa\_tor.doc).
\end{itemize}

\section{Inspection of Test Cases (by sampling)}
\begin{itemize}
\item \textbf{objectives:}
\begin{itemize}
\item check the representativeness of used programming language constructs for each considered coverage metrics;
\item check the compliance with the development standard.
\end{itemize}
\item \textbf{output:} QA inspection report (qa/DDMMYYYY/qa\_tor.doc, the same file used for the review of Tool Operational Requirements).
\end{itemize}

\section{Inspection of test execution results}
\begin{itemize}
\item \textbf{objectives:}
\begin{itemize}
\item Check the results of test execution
\item In the case tests failed, it is necessary to investigate whether the source of error is:
\begin{itemize}
\item A misbehaviour of the infrastructure used to run tests and compare actual results to expected results: in this case, the Qualification Team is in charge
of reporting and fixing the problem.
\item A bug in the \xcov implementation: in this case, the Development
Team is in charge of reporting and fixing the problem.
\item A reasonable limitation of the tool: in this case, the Qualification
Team is in charge of reporting and justifying the problem as part of \xcov known constraints and open problems.
\end{itemize}
\end{itemize}
\item \textbf{output:} QA inspection report (qa/DDMMYYYY/qa\_test\_execution.doc).
\end{itemize}

\section{Tool conformity review}
The conformity review takes in input a specific packaged and qualifiable release of \xcov.
\begin{itemize}
\item \textbf{objectives:}
\begin{itemize}
\item review the delivery file in the scope of the precise packaged and qualifiable release of \xcov
\item consistency within the Qualification Material;
\item finalize discussion on open problems (if any).
\end{itemize}
\item \textbf{output:} QA inspection report qa/<RELEASE>/DDMMYYYY/ qa\_conformity.doc).
\end{itemize}


\section{Responsibility}
The responsibility for the Quality Assurance Process belongs to the Quality Assurance Team.

\end{document}
